---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

Functions
```{r}
#functions

#MLR model function, no additional covariates
MCL<-function(data, blocks, blocksize, ref.grp, plot.param){
  colnames(data)<-c('marker', 'class')
  ds1sort<-data[order(data$marker),]
  #create equally sized groups of size BLOCKS to calculate DRL
  ds1sort$group<-c(rep(1:(blocks-1), each=blocksize), rep(blocks, times=blocksize-(blocks*blocksize-nrow(ds1sort))))
  #choose reference group
  ds1sort$categoryRef<-ifelse(ds1sort$group==ref.grp, 0, ds1sort$group)
  #fit model
  f1<-multinom(ds1sort$categoryRef ~ ds1sort$class, trace=FALSE)
  f2<-multinom(ds1sort$categoryRef ~ 1, trace=FALSE)
  an_f1f2_pval<-anova(f1, f2)[2,7]
  an_f1f2_TS<-anova(f1, f2)[2,6]
  
  
  #do wald test for parameters
  z <- summary(f1)$coefficients/summary(f1)$standard.errors
  # 2-tailed z test
  p <- (1 - pnorm(abs(z), 0, 1)) * 2
  p<-round(p, digits=4)
  #summarize model into table
  f1sum<-cbind(summary(f1)$coefficients, summary(f1)$standard.errors, exp(coef(f1)), z, p)
  colnames(f1sum)<-c("intercept", "beta", "SE(int)", "SE(beta)", "exp(int)", "exp(beta)", "int.Z", "beta.Z", "int.P", "beta.P")
  DLRJ<-(1+sum(exp(f1sum[,1]+f1sum[,2])))/(1+sum(f1sum[,5]))
  f1sum<-cbind(f1sum, DLRJ, an_f1f2_pval, an_f1f2_TS)
  
  colAUC(ds1sort$group, ds1sort$class, plotROC = plot.param)
  if(plot.param=='T'){
    abline(a=0, b=1)
  }
  return(f1sum)
}
#DLR calculated by contingency table, gives endpoints of all intervals
DLR<-function(data, blocks, blocksize){
  colnames(data)<-c('marker', 'class')
  ds1sort<-data[order(data$marker),]
  #create equally sized groups of size BLOCKS to calculate DRL
  ds1sort$group<-c(rep(1:(blocks-1), each=blocksize), rep(blocks, times=blocksize-(blocks*blocksize-nrow(ds1sort))))
  mat.all<-NULL
  min.all<-NULL
  max.all<-NULL
  for(i in 1:blocks){
    data_sub<-subset(ds1sort, ds1sort$group==i)
    min.range<-min(data_sub$marker)
    max.range<-max(data_sub$marker)
    min.all<-c(min.all, min.range)
    max.all<-c(max.all, max.range)
  }
  mat.all<-cbind(t(table(ds1sort$class, ds1sort$group)), min.all, max.all)
  N<-sum(mat.all[,1])
  P<-sum(mat.all[,2])
  DLR<-(mat.all[,2]/P)/(mat.all[,1]/N)
  mat.all<-cbind(mat.all[,1:2], DLR, mat.all[,3:4])
  colnames(mat.all)<-c("control", "case", "DLR", "min. of range", "max. of range")
  return(mat.all)
}
#Fit MLR model with a second categorical main effect
MCL_covar1_cat<-function(data, blocks, blocksize, ref.grp){
  colnames(data)<-c('marker', 'class', 'covar')
  ds1sort<-data[order(data$marker),]
  #create equally sized groups of size BLOCKS to calculate DRL
  ds1sort$group<-c(rep(1:(blocks-1), each=blocksize), rep(blocks, times=blocksize-(blocks*blocksize-nrow(ds1sort))))
  #choose reference group
  ds1sort$categoryRef<-ifelse(ds1sort$group==ref.grp, 0, ds1sort$group)
  #fit model
  f03<-multinom(ds1sort$categoryRef ~ ds1sort$class + factor(ds1sort$covar), trace=F)
  f_null<-multinom(ds1sort$categoryRef ~ ds1sort$class, trace=F)
  an_f1f2_pval<-anova(f03, f_null)[2,7]
  an_f1f2_TS<-anova(f03, f_null)[2,6]
  
  #do wald test for parameters
  z <- summary(f03)$coefficients/summary(f03)$standard.errors
  # 2-tailed z test
  p <- (1 - pnorm(abs(z), 0, 1)) * 2
  p<-round(p, digits=4)
  #summarize model into table
  f1sum<-cbind(summary(f03)$coefficients, summary(f03)$standard.errors, exp(coef(f03)), z, p)
  f1sum<-cbind(f1sum, an_f1f2_TS, an_f1f2_pval)
  return(f1sum)
}
#Cochran-Armitage test with 6 intervals
CA_test_10<-function(dat1, weights, blocksize){
  ds1sort<-dat1[order(dat1[,1]),]
  #create equally sized groups of size BLOCKS to calculate DRL
  ds1sort$group<-c(rep(1:(10-1), each=blocksize), rep(10, times=blocksize-(10*blocksize-nrow(ds1sort))))
  CochranArmitageTest(table(ds1sort[,2], ds1sort$group), alternative = 't')
  
  t1<-table(ds1sort[,2], ds1sort$group)
  R1<-sum(t1[1,])
  R2<-sum(t1[2,])
  Ci<-c(sum(t1[,1]),sum(t1[,2]),sum(t1[,3]),sum(t1[,4]),sum(t1[,5]),sum(t1[,6]),sum(t1[,7]),sum(t1[,8]),sum(t1[,9]),sum(t1[,10]))
  
  weights<-weights
  
  CA_TS<-sum(weights*(t1[1,]*R2-t1[2,]*R1))
  
  V1<-(weights[1]*Ci[1])*(weights[2]*Ci[2] + weights[3]*Ci[3] + weights[4]*Ci[4] + weights[5]*Ci[5] +
                            weights[6]*Ci[6] + weights[7]*Ci[7] + weights[8]*Ci[8] + weights[9]*Ci[9] + weights[10]*Ci[10])
  V2<-(weights[2]*Ci[2])*(weights[3]*Ci[3] + weights[4]*Ci[4] + weights[5]*Ci[5] + weights[6]*Ci[6] +
                            weights[7]*Ci[7] + weights[8]*Ci[8] + weights[9]*Ci[9] + weights[10]*Ci[10])
  V3<-(weights[3]*Ci[3])*(weights[4]*Ci[4] + weights[5]*Ci[5] + weights[6]*Ci[6] + weights[7]*Ci[7] +
                            weights[8]*Ci[8] + weights[9]*Ci[9] + weights[10]*Ci[10])
  V4<-(weights[4]*Ci[4])*(weights[5]*Ci[5] + weights[6]*Ci[6] + weights[7]*Ci[7] + weights[8]*Ci[8] +
                            weights[9]*Ci[9] + weights[10]*Ci[10])
  V5<-(weights[5]*Ci[5])*(weights[6]*Ci[6] + weights[7]*Ci[7] + weights[8]*Ci[8] + weights[9]*Ci[9] + weights[10]*Ci[10])
  V6<-(weights[6]*Ci[6])*(weights[7]*Ci[7] + weights[8]*Ci[8] + weights[9]*Ci[9] + weights[10]*Ci[10])
  V7<-(weights[7]*Ci[7])*(weights[8]*Ci[8] + weights[9]*Ci[9] + weights[10]*Ci[10])
  V8<-(weights[8]*Ci[8])*(weights[9]*Ci[9] + weights[10]*Ci[10])
  V9<-(weights[9]*Ci[9])*(weights[10]*Ci[10])
  Vi<-sum(V1, V2, V3, V4, V5, V6, V7, V8, V9)
  
  CA_Var<-((R1*R2)/nrow(dat1))*(sum(((weights**2)*Ci)*(sum(t1[,])-Ci)) - (2*Vi))
  
  q<-abs(CA_TS/sqrt(CA_Var))
  
  #calculate sigma matrix
  N<-length(ds1sort$group)
  c_vec<-colSums(t1)
  sigma_mat<-diag(c_vec*(N-c_vec))
  sigma_mat[1,2:10]<--c_vec[1]*c_vec[2:10]
  sigma_mat[2:10,1]<--c_vec[1]*c_vec[2:10]
  
  sigma_mat[2,3:10]<--c_vec[2]*c_vec[3:10]
  sigma_mat[3:10,2]<--c_vec[2]*c_vec[3:10]
  
  sigma_mat[3,4:10]<--c_vec[3]*c_vec[4:10]
  sigma_mat[4:10,3]<--c_vec[3]*c_vec[4:10]
  
  sigma_mat[4,5:10]<--c_vec[4]*c_vec[5:10]
  sigma_mat[5:10,4]<--c_vec[4]*c_vec[5:10]
  
  sigma_mat[5,6:10]<--c_vec[5]*c_vec[6:10]
  sigma_mat[6:10,5]<--c_vec[5]*c_vec[6:10]
  
  sigma_mat[6,7:10]<--c_vec[6]*c_vec[7:10]
  sigma_mat[7:10,6]<--c_vec[6]*c_vec[7:10]
  
  sigma_mat[7,8:10]<--c_vec[7]*c_vec[8:10]
  sigma_mat[8:10,7]<--c_vec[7]*c_vec[8:10]
  
  sigma_mat[8,9:10]<--c_vec[8]*c_vec[9:10]
  sigma_mat[9:10,8]<--c_vec[8]*c_vec[9:10]
  
  sigma_mat[9,10]<--c_vec[9]*c_vec[10]
  sigma_mat[10,9]<--c_vec[9]*c_vec[10]
  
  sigma_mat<-((rowSums(t1)[1]*rowSums(t1)[2])/N)*sigma_mat
  
  p.val<-pnorm(q, lower.tail = F)*2
  
  return(list(CA_TS, p.val, sigma_mat))
}


library(readxl); library(mcp.project); library(caTools); library(nnet);library(msm);library(pROC);library(xtable);library(plyr)
```


Read in data and create outcome
```{r}
dov<-read_excel("VAN_HGSOC.xlsx")
table(dov$Stage)
dov<-subset(dov, dov$Stage!=8)
dov$Stage<-ifelse(dov$Stage==1, 0, 1) #recode stage
```

LR test to identify significant genes
```{r}
dov<-data.frame(dov)
#apply LRT to all genes
an_pv_all<-NULL
for(i in 14:526){ #markers start at column 14
  fit1<-MCL(dov[,c(i,8)], 10, 40, 10, 'F')[1,12]
  an_pv_all<-c(an_pv_all, fit1)
}

#get adjusted p-values
adj_pval<-fdr(an_pv_all, method='BH', q=.05)
table(adj_pval$Pvals$rejected) #118 rejections
sig_genes<-which(adj_pval$Pvals$rejected==T)+13 #gives column number of dov for sig genes
```


CA test to categorize significant genes
```{r}
CAtest_all<-NULL
TS_diff_all<-NULL
pval_all<-NULL
for(i in sig_genes){
  
  mean_cont<-mean(dov[which(dov$Stage==0),i])
  mean_case<-mean(dov[which(dov$Stage==1),i])
  
  if(mean_cont>mean_case){
    dov$Stage2<-ifelse(dov$Stage==1, 0, 1)
  } else{
    dov$Stage2<-dov$Stage
  }
  
  CAtest_trad<-(CA_test_10(data.frame(dov[,i], dov$Stage2), 10:1, 40))
  CAtest_nt<-(CA_test_10(data.frame(dov[,i], dov$Stage2), c(9,7,5,3,1,0,2,4,6,8), 40))
  
  difftruth<-CAtest_trad[[1]]+CAtest_nt[[1]]
  
  weight_trad<-matrix(data=c(10:1), nrow = 10)
  weight_nt<-c(9,7,5,3,1,0,2,4,6,8)
  var_trad<-t(weight_trad)%*%CAtest_trad[[3]]%*%weight_trad
  var_nt<-t(weight_nt)%*%CAtest_nt[[3]]%*%weight_nt
  covar_T<-t(weight_trad)%*%CAtest_trad[[3]]%*%weight_nt
  
  (TS_diff<-difftruth/sqrt(var_nt + var_trad + (2*covar_T)))
  TS_diff_all<-c(TS_diff_all, TS_diff)
  
  if(TS_diff<0){
    p_val1<-pnorm(TS_diff, lower.tail = T)*2
  } else{
    p_val1<-pnorm(TS_diff, lower.tail = F)*2
  }
  
  pval_all<-c(pval_all, p_val1)
}


trad_genes<-sig_genes[which(TS_diff_all > 0 & pval_all < 0.05)] #column # of dov with traditional sig. genes
nt_genes<-sig_genes[which(TS_diff_all < 0 & pval_all < 0.05)] #column # of dov with nontraditional sig. genes

length(sig_genes)
length(nt_genes) + length(trad_genes)
#21 indeterminate
#questions:
#1. what do the 100 trad look like
#2. What do the 15 NT look like
#3 what do the 21 indeterminate look like? 
`%!in%` <- Negate(`%in%`)
indet <- sig_genes[which(sig_genes %!in% c(trad_genes, nt_genes))]

par(mfrow=c(3,3))
for(i in indet){
  colAUC(as.numeric(bin(dov[,i], nbins=10, method='content', labels=c(1:10))), dov[,8], plotROC = T)
  abline(a=0,b=1)
}

par(mfrow=c(3,3))
for(i in nt_genes){
  colAUC(as.numeric(bin(dov[,i], nbins=10, method='content', labels=c(1:10))), dov[,8], plotROC = T)
  abline(a=0,b=1)
}

par(mfrow=c(3,3))
for(i in trad_genes){
  colAUC(as.numeric(bin(dov[,i], nbins=10, method='content', labels=c(1:10))), dov[,8], plotROC = T)
  abline(a=0,b=1)
}
```

AUC test to identify informative markers
```{r}
auc_test_all<-NULL
for(i in 14:526){
    auc_test<-wilcox.test(subset(dov, dov$Stage==1)[,i], subset(dov, dov$Stage==0)[,i])$`p.val`
    auc_test_all<-c(auc_test_all, auc_test)
}

#adjust for multiple testing
table(ifelse(auc_test_all <0.05, 1, 0)) #213 rejections
adj_pval_auc<-fdr(auc_test_all, method='BH', q=.05)
table(adj_pval_auc$Pvals$rejected) #166  rejections
sig_genes_auc<-which(adj_pval_auc$Pvals$rejected==T)+13 #column #s of significant genes by AUC
```

Test for trend for AUC significant genes
```{r}
CAtest_all<-NULL
TS_diff_all<-NULL
pval_all<-NULL
for(i in sig_genes_auc){
  
  mean_cont<-mean(dov[which(dov$Stage==0),i])
  mean_case<-mean(dov[which(dov$Stage==1),i])
  
  if(mean_cont>mean_case){
    dov$Stage2<-ifelse(dov$Stage==1, 0, 1)
  } else{
    dov$Stage2<-dov$Stage
  }
  
  CAtest_trad<-(CA_test_10(data.frame(dov[,i], dov$Stage2), 10:1, 40))
  CAtest_nt<-(CA_test_10(data.frame(dov[,i], dov$Stage2), c(9,7,5,3,1,0,2,4,6,8), 40))
  
  difftruth<-CAtest_trad[[1]]+CAtest_nt[[1]]
  
  weight_trad<-matrix(data=c(10:1), nrow = 10)
  weight_nt<-c(9,7,5,3,1,0,2,4,6,8)
  var_trad<-t(weight_trad)%*%CAtest_trad[[3]]%*%weight_trad
  var_nt<-t(weight_nt)%*%CAtest_nt[[3]]%*%weight_nt
  covar_T<-t(weight_trad)%*%CAtest_trad[[3]]%*%weight_nt
  
  (TS_diff<-difftruth/sqrt(var_nt + var_trad + (2*covar_T)))
  TS_diff_all<-c(TS_diff_all, TS_diff)
  
  if(TS_diff<0){
    p_val1<-pnorm(TS_diff, lower.tail = T)*2
  } else{
    p_val1<-pnorm(TS_diff, lower.tail = F)*2
  }
  
  pval_all<-c(pval_all, p_val1)
}


trad_genes_auc<-sig_genes_auc[which(TS_diff_all > 0 & pval_all < 0.05)] #column # of dov with traditional sig. genes
nt_genes_auc<-sig_genes_auc[which(TS_diff_all < 0 & pval_all < 0.05)] #column # of dov with nontraditional sig. genes

indet_auc <- sig_genes_auc[which(sig_genes_auc %!in% c(trad_genes_auc))]
```

Compare LRT rejections to AUC rejections
```{r}
table(adj_pval_auc$Pvals$rejected, adj_pval$Pvals$rejected)
#auc as rows, LRT as column
#they agreed on rejecting 115
#21 that LRT thought were important that AUC didnt
#51 that AUC thought were important that AUC didnt
#326 they agreed on failing to reject

#look at the 21 AUC missed
discord_pairs<-ifelse(adj_pval_auc$Pvals$rejected==F & adj_pval$Pvals$rejected==T, 1, 0)
auc_miss<-which(discord_pairs==1)+13 #column numbers of the biomarkers the AUC test missed compared to the LR test
```


Plot the 21 AUC missed
```{r}
for(i in auc_miss){
  colAUC(dov[,i], dov[,8], plotROC = T)
  abline(a=0,b=1)
}

#11 out of the 26 were on the NT list
#9 out of the 26 were on the indet
# 6 out of the 26 were on the trad list
sum(auc_miss %in% trad_genes)

#The 11 account for what percentage of the total number of NTBs? 100%
sum(auc_miss %in% nt_genes)/length(nt_genes)
```

Test for length of ROC curve to identify informative biomarkers
```{r}
length_cont<-length(dov[which(dov$Stage==0),8])
length_case<-length(dov[which(dov$Stage==1),8])

pval_all_length<-NULL
for(i in 14:526){
  mu0<-mean(dov[which(dov$Stage==0),i])
  mu1<-mean(dov[which(dov$Stage==1),i])
  sd0<-sd(dov[which(dov$Stage==0),i])
  sd1<-sd(dov[which(dov$Stage==1),i])
  
  #length function
  l1<-function(x){
    sqrt(dnorm(x, mu0, sd0)^2 + (dnorm(x, mu1, sd1))^2) #assumes binormal model for markers
  }
  
  l1_est<-integrate(l1, lower=min(dov[,i]), upper = max(dov[,i]))
  
  
  #inference procedure
  #note that the null hypothesis corresponds to a noninformative biomarker than follows the 45-degree line
  Z<-((sqrt(2)*4*length_cont*length_case)/(length_cont + length_case))*(l1_est$value - sqrt(2))
  
  pval_all_length<-c(pval_all_length, pchisq(Z, 2, lower.tail = F))

}

#adjust for multiple testing
table(ifelse(pval_all_length <0.05, 1, 0)) #140 rejections
adj_pval_length<-fdr(pval_all_length, method='BH', q=.05)
table(adj_pval_length$Pvals$rejected) #73  rejections
sig_genes_length<-which(adj_pval_length$Pvals$rejected==T)+13 #column #s of significant genes by ROC length
```

Test for trend for length significant genes
```{r}
CAtest_all<-NULL
TS_diff_all<-NULL
pval_all<-NULL
for(i in sig_genes_length){
  
  mean_cont<-mean(dov[which(dov$Stage==0),i])
  mean_case<-mean(dov[which(dov$Stage==1),i])
  
  if(mean_cont>mean_case){
    dov$Stage2<-ifelse(dov$Stage==1, 0, 1)
  } else{
    dov$Stage2<-dov$Stage
  }
  
  CAtest_trad<-(CA_test_10(data.frame(dov[,i], dov$Stage2), 10:1, 40))
  CAtest_nt<-(CA_test_10(data.frame(dov[,i], dov$Stage2), c(9,7,5,3,1,0,2,4,6,8), 40))
  
  difftruth<-CAtest_trad[[1]]+CAtest_nt[[1]]
  
  weight_trad<-matrix(data=c(10:1), nrow = 10)
  weight_nt<-c(9,7,5,3,1,0,2,4,6,8)
  var_trad<-t(weight_trad)%*%CAtest_trad[[3]]%*%weight_trad
  var_nt<-t(weight_nt)%*%CAtest_nt[[3]]%*%weight_nt
  covar_T<-t(weight_trad)%*%CAtest_trad[[3]]%*%weight_nt
  
  (TS_diff<-difftruth/sqrt(var_nt + var_trad + (2*covar_T)))
  TS_diff_all<-c(TS_diff_all, TS_diff)
  
  if(TS_diff<0){
    p_val1<-pnorm(TS_diff, lower.tail = T)*2
  } else{
    p_val1<-pnorm(TS_diff, lower.tail = F)*2
  }
  
  pval_all<-c(pval_all, p_val1)
}


trad_genes_length<-sig_genes_length[which(TS_diff_all > 0 & pval_all < 0.05)] #column # of dov with traditional sig. genes
nt_genes_length<-sig_genes_length[which(TS_diff_all < 0 & pval_all < 0.05)] #column # of dov with nontraditional sig. genes

indet_length <- sig_genes_length[which(sig_genes_length %!in% c(trad_genes_length, nt_genes_length))]

par(mfrow=c(3,3))
for(i in indet_length){
  colAUC(as.numeric(bin(dov[,i], nbins=10, method='content', labels=c(1:10))), dov[,8], plotROC = T)
  abline(a=0,b=1)
}

par(mfrow=c(3,3))
for(i in nt_genes_length){
  colAUC(as.numeric(bin(dov[,i], nbins=10, method='content', labels=c(1:10))), dov[,8], plotROC = T)
  abline(a=0,b=1)
}

par(mfrow=c(3,3))
for(i in trad_genes_length){
  colAUC(as.numeric(bin(dov[,i], nbins=10, method='content', labels=c(1:10))), dov[,8], plotROC = T)
  abline(a=0,b=1)
}
```

Compare LRT rejections to length rejections
```{r}
table(adj_pval_length$Pvals$rejected, adj_pval$Pvals$rejected)
#length as rows, LRT as column
#they agreed on rejecting 58
#60 that LRT thought were important that length didnt
#15 that length thought were important that LRT didnt
#380 they agreed on failing to reject

#look at the 60 length missed
discord_pairs<-ifelse(adj_pval_length$Pvals$rejected==F & adj_pval$Pvals$rejected==T, 1, 0)
length_miss<-which(discord_pairs==1)+13 #column numbers of the biomarkers the AUC test missed compared to the LR test
#look at the 15 LRT missed
discord_pairs2<-ifelse(adj_pval_length$Pvals$rejected==T & adj_pval$Pvals$rejected==F, 1, 0)
length_miss2<-which(discord_pairs2==1)+13
```

Plot the 60 length missed
```{r}
par(mfrow=c(3,3))
for(i in length_miss){
  colAUC(dov[,i], dov[,8], plotROC = T)
  abline(a=0,b=1)
}

#37 out of the 60 were on the trad list
#10 out of the 60 were on the ntrad list
#13 out of the 60 were on the nt list
sum(length_miss %in% nt_genes)
length_miss_nt<-length_miss[which(length_miss %in% nt_genes)]
par(mfrow=c(3,3))
rov_miss_all<-NULL
auc_tradmiss_all<-NULL
for(i in length_miss_nt){
  auc_tradmiss<-colAUC(dov[,i], dov[,8], plotROC = F)
  #abline(a=0,b=1)
  auc_tradmiss_all<-c(auc_tradmiss_all, auc_tradmiss)
  
  rov<-sd(dov[which(dov$Stage==1),i])/sd(dov[which(dov$Stage==0),i])
  rov<-ifelse(rov<1, 1/rov, rov)
  rov_miss_all<-c(rov_miss_all, rov)
}
summary(rov_miss_all)
summary(auc_tradmiss_all)


#plot the 15 that LRT missed but length picked up
par(mfrow=c(4,4))
for(i in length_miss2){
  colAUC(dov[,i], dov[,8], plotROC = T)
  abline(a=0,b=1)
}

sum(length_miss2 %in% indet_length)
#of the 15, 9 were in the trad_length group, 1 in the nt_length group, and 5 in the indet_length group
miss2_trad<-length_miss2[which(length_miss2 %in% trad_genes_length)]
  auc_miss_all<-NULL
  for(i in miss2_trad){
    auc_miss_all<-c(auc_miss_all, colAUC(dov[,i], dov[,8], plotROC = F))
  }
  summary(auc_miss_all)

  miss2_nt<-length_miss2[which(length_miss2 %in% nt_genes_length)]
  rov_miss_all<-NULL
  for(i in miss2_nt){
    rov<-sd(dov[which(dov$Stage==1),i])/sd(dov[which(dov$Stage==0),i])
    rov<-ifelse(rov<1, 1/rov, rov)
    rov_miss_all<-c(rov_miss_all, rov)
  }
  summary(rov_miss_all)
miss2_indet<-length_miss2[which(length_miss2 %in% indet_length)]
```

Find most significant traditional and nontraditional
```{r}
#most sig. nontrad by LRT
nt_genes[which(cbind(nt_genes, an_pv_all[nt_genes - 13])[,2] == min(cbind(nt_genes, an_pv_all[nt_genes - 13])[,2]))]
colnames(dov)[377]

#most sig. trad by LRT
trad_genes[which(cbind(trad_genes, an_pv_all[trad_genes - 13])[,2] == min(cbind(trad_genes, an_pv_all[trad_genes - 13])[,2]))]
colnames(dov)[442]
```


Plots of most significant traditional and nontraditional
```{r}
par(mfrow=c(2,2))
fit1<-MCL(dov[,c(442,8)], 10, 40, 10, 'F')
DLR_val<-c(fit1[,6]/fit1[1,11], 1/fit1[1,11])

bins<-DLR(dov[,c(442,8)], 10, 40)
par(mfrow=c(2,2))
plot(seq(bins[1,4], bins[1,5], length.out = 2), rep(DLR_val[1], times=2), type='l', 
     ylim=c(0,5.5), xlim=c(bins[1,4], bins[10,5]), ylab = 'DLRj', xlab = 'log-standardized marker', main='DLR Curve', sub='Gene: PCDH9', xaxt='n', cex.lab=1.2, cex.axis=1.2)
abline(h=1, lty=2, col='grey')
lines(seq(bins[2,4], bins[2,5], length.out = 2), rep(DLR_val[2], times=2))
lines(seq(bins[3,4], bins[3,5], length.out = 2), rep(DLR_val[3], times=2))
lines(seq(bins[4,4], bins[4,5], length.out = 2), rep(DLR_val[4], times=2))
lines(seq(bins[5,4], bins[5,5], length.out = 2), rep(DLR_val[5], times=2))
lines(seq(bins[6,4], bins[6,5], length.out = 2), rep(DLR_val[6], times=2))
lines(seq(bins[7,4], bins[7,5], length.out = 2), rep(DLR_val[7], times=2))
lines(seq(bins[8,4], bins[8,5], length.out = 2), rep(DLR_val[8], times=2))
lines(seq(bins[9,4], bins[9,5], length.out = 2), rep(DLR_val[9], times=2))
lines(seq(bins[10,4], bins[10,5], length.out = 2), rep(DLR_val[10], times=2))
axis(1, at=c(bins[,4], bins[10,5]), labels=round(c(bins[,4], bins[10,5]), digits=1), cex.axis=1)

ds1sort<-dov[order(dov$G429),]
ds1sort$group<-c(rep(1:(10-1), each=40), rep(10, times=40-(10*40-nrow(ds1sort))))
fit1<-glm(ds1sort$Stage ~ ds1sort$group, family = binomial)
pred429<-predict(fit1)
roc429<-roc(ds1sort$Stage, pred429, plot = F)
ci.sp.obj <- ci.sp(roc429, sensitivities=roc429$sensitivities, boot.n=100)
plot(roc429, ylim=c(0,1), main='ROC Curve', cex.lab=1.2, cex.axis=1.2) # restart a new plot
plot(ci.sp.obj, type="shape", col="grey", ylim=c(0,1))
abline(a=1,b=-1, lwd=2)


fit1<-MCL(dov[,c(377,8)], 10, 40, 10, 'F')
DLR_val<-c(fit1[,6]/fit1[1,11], 1/fit1[1,11])

bins<-DLR(dov[,c(377,8)], 10, 40)
plot(seq(bins[1,4], bins[1,5], length.out = 2), rep(DLR_val[1], times=2), type='l',
     ylim=c(0,5.5), xlim=c(bins[1,4], bins[10,5]), ylab = 'DLRj', xlab = 'log-standardized marker', main='DLR Curve', sub='Gene: XRN2', xaxt='n', cex.lab=1.2, cex.axis=1.2)
abline(h=1, lty=2, col='grey')
lines(seq(bins[2,4], bins[2,5], length.out = 2), rep(DLR_val[2], times=2))
lines(seq(bins[3,4], bins[3,5], length.out = 2), rep(DLR_val[3], times=2))
lines(seq(bins[4,4], bins[4,5], length.out = 2), rep(DLR_val[4], times=2))
lines(seq(bins[5,4], bins[5,5], length.out = 2), rep(DLR_val[5], times=2))
lines(seq(bins[6,4], bins[6,5], length.out = 2), rep(DLR_val[6], times=2))
lines(seq(bins[7,4], bins[7,5], length.out = 2), rep(DLR_val[7], times=2))
lines(seq(bins[8,4], bins[8,5], length.out = 2), rep(DLR_val[8], times=2))
lines(seq(bins[9,4], bins[9,5], length.out = 2), rep(DLR_val[9], times=2))
lines(seq(bins[10,4], bins[10,5], length.out = 2), rep(DLR_val[10], times=2))
axis(1, at=c(bins[,4], bins[10,5]), labels=round(c(bins[,4], bins[10,5]), digits=1), cex.axis=1)

ds1sort<-dov[order(dov$G364),]
ds1sort$group<-c(rep(1:(10-1), each=40), rep(10, times=40-(10*40-nrow(ds1sort))))
fit1<-glm(ds1sort$Stage ~ ds1sort$group, family = binomial)
pred364<-predict(fit1)
roc364<-roc(ds1sort$Stage, pred364, plot = F)
ci.sp.obj <- ci.sp(roc364, sensitivities=roc364$sensitivities, boot.n=100)
plot(roc364, ylim=c(0,1), main='ROC Curve', cex.lab=1.2, cex.axis=1.2) 
plot(ci.sp.obj, type="shape", col="grey", ylim=c(0,1))
abline(a=1,b=-1, lwd=2)
```


Obtaining post-test probabilities
```{r}
#fit model with stage and age
fit6<-MCL_covar1_cat(data.frame(cbind(as.numeric(dov[,447]), dov$Stage, dov$quartile_age)), 10, 40, 10)

447 %in% trad_genes

#conduct chi-square test of independence
data<-data.frame(cbind(dov$G434, dov$Stage, dov$quartile_age))
colnames(data)<-c('marker', 'class', 'age')
ds1sort<-data[order(data$marker),]
#create equally sized groups of size BLOCKS to calculate DRL
ds1sort$group<-c(rep(1:(10-1), each=40), rep(10, times=40-(10*40-nrow(ds1sort))))
table(ds1sort$group, ds1sort$age)
chisq.test(table(ds1sort$group, ds1sort$age)) #marker group and age are not independent

#calculate DLR for each combination of covariates
DLRS1A1<-function(j){exp(fit6[j,2])*((1+sum(exp(fit6[,1])))/(1+sum(exp(fit6[,1]+fit6[,2]))))}
DLRS1A2<-function(j){exp(fit6[j,2])*((1+sum(exp(fit6[,1]+fit6[,3])))/(1+sum(exp(fit6[,1]+fit6[,2]+fit6[,3]))))}
DLRS1A3<-function(j){exp(fit6[j,2])*((1+sum(exp(fit6[,1]+fit6[,4])))/(1+sum(exp(fit6[,1]+fit6[,2]+fit6[,4]))))}
DLRS1A4<-function(j){exp(fit6[j,2])*((1+sum(exp(fit6[,1]+fit6[,5])))/(1+sum(exp(fit6[,1]+fit6[,2]+fit6[,5]))))}

DLRS1A10<-((1+sum(exp(fit6[,1])))/(1+sum(exp(fit6[,1]+fit6[,2]))))
DLRS1A20<-((1+sum(exp(fit6[,1]+fit6[,3])))/(1+sum(exp(fit6[,1]+fit6[,2]+fit6[,3]))))
DLRS1A30<-((1+sum(exp(fit6[,1]+fit6[,4])))/(1+sum(exp(fit6[,1]+fit6[,2]+fit6[,4]))))
DLRS1A40<-((1+sum(exp(fit6[,1]+fit6[,5])))/(1+sum(exp(fit6[,1]+fit6[,2]+fit6[,5]))))

#plot DLR function adjusted for stage for each age category
#age group 1
DLRS1A1_val<-round(c(sapply(X=c(1:9), FUN=DLRS1A1), DLRS1A10), digits=4)
#age group 2
DLRS1A2_val<-round(c(sapply(X=c(1:9), FUN=DLRS1A2), DLRS1A20), digits=4)
#age group 3
DLRS1A3_val<-round(c(sapply(X=c(1:9), FUN=DLRS1A3), DLRS1A30), digits=4)
#age group 4
DLRS1A4_val<-round(c(sapply(X=c(1:9), FUN=DLRS1A4), DLRS1A40), digits=4)

#calculate age DLR
t3<-table(dov$Stage, dov$quartile_age)
S1A1<-(t3[2,1]/sum(t3[2,]))/(t3[1,1]/sum(t3[1,]))
S1A2<-(t3[2,2]/sum(t3[2,]))/(t3[1,2]/sum(t3[1,]))
S1A3<-(t3[2,3]/sum(t3[2,]))/(t3[1,3]/sum(t3[1,]))
S1A4<-(t3[2,4]/sum(t3[2,]))/(t3[1,4]/sum(t3[1,]))

#apply serial testing strategy (age applied first, then G9 adjusted for age)

tab1<-cbind(rep(.25, times=40), rep(c('<53', '53-59', '60-66', '>66'), each=10), rep(round(c(S1A1, S1A2, S1A3, S1A4), digits=2), each=10), rep(1:10, times=4), round(bins[,3], digits=2), round(c(DLRS1A1_val, DLRS1A2_val, DLRS1A3_val, DLRS1A4_val), digits=2))

post_odds<-as.numeric(tab1[,1])/(1-as.numeric(tab1[,1]))*as.numeric(tab1[,3])*as.numeric(tab1[,6])
post_prob<-post_odds/(1+post_odds)
tab1<-cbind(tab1, round(post_prob, digits=2))

colnames(tab1)<-c('Pre Prob', 'Age','DLR(Age)','Marker interval', 'DLR(G90)','DLR(G90|Age)', 'Post Prob')

#add row if tests had been applied sequentially
post_odds_seq<-round(c(as.numeric(tab1[1,1])/(1-as.numeric(tab1[1,1]))*S1A1*bins[,3], as.numeric(tab1[1,1])/(1-as.numeric(tab1[1,1]))*S1A2*bins[,3], as.numeric(tab1[1,1])/(1-as.numeric(tab1[1,1]))*S1A3*bins[,3], as.numeric(tab1[1,1])/(1-as.numeric(tab1[1,1]))*S1A4*bins[,3]), digits=2)

post_prob_seq<-post_odds_seq/(1+post_odds_seq)

tab2<-cbind(tab1, round(post_prob_seq, digits=2))
colnames(tab2)<-c('Pre Prob', 'Age','DLR(Age)','Marker interval', 'DLR(G90)','DLR(G90|Age)', 'covariate-adjusted p1', 'indep. p1')

options(xtable.comment = FALSE)
print(xtable(tab2, label = NULL, caption = 'Post-test probabilities'), include.rownames=FALSE)
```

